Low Dimensional Embedding of Climate Data for Radio Astronomical Site Testing in the Colombian Andes

### 2017-05-11 mnras/mnras_template.tex
mnras

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{nicefrac} 
\usepackage{epstopdf}
\usepackage{booktabs}

@@

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
keyword1 -- keyword2 -- keyword3
\end{keywords}

@@

In this paper we seek to find plausible locations for a high-mountain mm-wave astronomical observatory in the northern Andes of Colombia, by analizing 30 years of climate data from 2046 weather stations. By low-dimensionally embedding the data, we are able to group together and correlate climate behavior indicative of daytime dry, clear-sky weather conditions at elevations higher than 2000 masl.  A repository with all our code and the full dataset is available at \texttt{https://github.com/saint-germain/ideam}.\\

From our shortlist of 79 stations that met our criteria, we selected 70 stations located  in the Cundinamarca-Boyac\'a \emph{altiplano} region of Colombia, which is a high-mountain basin with an average altitude of 2600 masl. Weather stations in our shortlist but outside this region are too sparsely located to suggest a candidate region of interest. \\

@@

We summarize the information about two regions of interest (Valle de Ubat\'e and Valle del Sol) in Table \ref{tabfin}, selected based on a Bayesian probabilistic quality index. Most of the stations in this table are located at elevations above 2600 masl and below 3000 masl. We validated our results using satellite measurements of the upper bound for precipitable water vapor \citep{suen2014,suen2016} for those two regions, whis is reported to be $\sim15$ mm. This means that it is not implausible to find a site with an even lower water vapor column in one of the regions of interest identified here. Some mountains nearby Valle de Ubat\'e and Valle del Sol  have altitudes of up to 3400 masl, and therefore warrant further measurements using radio sondes and radiometers. \\

>>>>

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{nicefrac} 
\usepackage{booktabs}

@@

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
atmospheric effects -- methods: data analysis -- methods: statistical -- site testing
\end{keywords}

@@

In this paper we find plausible locations for a high-mountain mm-wave astronomical observatory in the northern Andes of Colombia, by analizing 30 years of climate data from 2046 weather stations. By low-dimensionally embedding the data, we are able to group together and correlate climate behavior indicative of daytime dry, clear-sky weather conditions.  A repository with all our code and the full dataset is available at \texttt{https://github.com/saint-germain/ideam}.\\

From our shortlist of 79 stations at elevations higher than 2000 masl that met our criteria, we selected 70 stations located  in the Cundinamarca-Boyac\'a \emph{altiplano} region of Colombia, which is a high-mountain basin with an average altitude of 2600 masl. Weather stations in our shortlist but outside this region are too sparsely located to suggest a candidate region of interest. \\

@@

We summarize the information about two regions of interest (Valle de Ubat\'e and Valle del Sol) in Table \ref{tabfin}, selected based on a Bayesian probabilistic quality index. Most of the stations in this table are located at elevations above 2600 masl and below 3000 masl. We validated our results using satellite measurements of the upper bound for mean precipitable water vapor \citep{suen2014,suen2016} for those two regions, which is reported to be $\sim15$ mm in $\sim20$ km resolution maps. This means that it is not implausible to find a site with an even lower water vapor column in one of the regions of interest identified here. Some mountains nearby Valle de Ubat\'e and Valle del Sol  have altitudes of up to 3400 masl, and therefore warrant further measurements using radio sondes and radiometers. \\
